Global search algorithm with local tuning for solving mixed-integer global optimization problems \thanks{This study was supported by the .}

### 2021-02-13 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex

\begin{document}
%
\title{An Approach for Simultaneous Finding of Multiple Effective Decisions in Multi-objective Optimization Problems
%with non-convex constraints 
\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.0729-2020-0055, and by the Research and Education Mathematical Center, project no. 075-02-2020-1483/1.}}
%
\titlerunning{Finding of Multiple Effective Decisions in MOO Problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
Victor Gergel\orcidID{0000-0002-4013-2329} \and
Vladimir Grishagin \and
Evgeniy Kozinov\orcidID{0000-0001-6776-0096} 
}
%
\authorrunning{K. Barkalov, V. Gergel, V. Grishagin, E. Kozinov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru,\{gergel,vagris\}@unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This paper considers computationally intensive multi-objective optimization problems which require computing multiple Pareto-optimal decisions. It is also assumed that efficiency criteria may be multiextremal, and the cost of calculating function values may be quite high. The proposed approach is based on the reduction of multi-objective optimization problems to one-dimensional global optimization problems that can be solved using effective information-statistical algorithms of global search. One key innovation of the developed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained. Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multi-objective optimization problems. Computational experiments confirm the potential of the proposed approach.

@@

\textbf{1. Scalarization of multiple objective functions.} In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
\begin{equation}
\label{eq:04}
\min {\varphi(y)} = F(\lambda, y), y \in D,
\end{equation}
where $F$ is a scalar multiextremal function generated as a result of scalarization of objective functions $f_i$, $1 \leq i \leq s$, $\lambda$ is the vector of parameters of the applied convolution of functions, and $D$ is the search domain from (\ref{eq:01}). In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the values of the objective functions of the specified reference decision $y^0 \in D$. In this case, a possible statement of the statement scalar optimization problem can have the form 
\begin{equation}
\label{eq:05}
\min F(\lambda,y)=\min 1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
\end{equation}
where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately. Without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set 
\begin{equation}
\label{eq:06}
\lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
\end{equation}

@@

\textbf{1. The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements. Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}). These functions $F(\lambda, y)$ for different values of the coefficients $\lambda \in \Lambda$ can be optimized sequentially. This determines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the efficient decisions found.

>>>>

\begin{document}
%
\title{An Approach for Simultaneous Finding of Multiple Effective Decisions in Multi-objective Optimization Problems
%with non-convex constraints 
\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.0729-2020-0055, and by the Research and Education Mathematical Center, project no. 075-02-2020-1483/1.}}
%
\titlerunning{Finding of Multiple Effective Decisions in MOO Problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
Victor Gergel\orcidID{0000-0002-4013-2329} \and
Vladimir Grishagin \and
Evgeniy Kozinov\orcidID{0000-0001-6776-0096} 
}
%
\authorrunning{K. Barkalov, V. Gergel, V. Grishagin, E. Kozinov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru,\{gergel,vagris\}@unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This paper considers computationally intensive multi-objective optimization problems which require computing multiple Pareto-optimal decisions. It is also assumed that efficiency criteria may be multiextremal, and the cost of calculating function values may be quite high. The proposed approach is based on the reduction of multi-objective optimization problems to one-dimensional global optimization problems that can be solved using efficient information-statistical algorithms of global search. One key innovation of the developed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained. Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multi-objective optimization problems. Computational experiments confirm the potential of the proposed approach.

@@

\textbf{1. Scalarization of multiple objective functions.} In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
\begin{equation}
\label{eq:04}
\min {\varphi(y)} = F(\lambda, y), y \in D,
\end{equation}
where $F$ is a scalar multiextremal function generated as a result of scalarization of objective functions $f_i$, $1 \leq i \leq s$, $\lambda$ is the vector of parameters of the applied convolution of functions, and $D$ is the search domain from (\ref{eq:01}). In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the values of the objective functions of the specified reference decision $y^0 \in D$. In this case, a possible statement of scalar optimization problem can have the form 
\begin{equation}
\label{eq:05}
\min F(\lambda,y)=\min 1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
\end{equation}
where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately. Without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set 
\begin{equation}
\label{eq:06}
\lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
\end{equation}

@@

\textbf{1. The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements. Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}). These functions $F(\lambda, y)$ can be optimized sequentially. This determines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the efficient decisions found.
